---
title: "Formative Assessment 7"
author: "Ramilo, Zion John Yousef T."
output: github_document
---
```{r setup, include=FALSE}
library(datasets)
library(tidyverse)
library(car)
library(Hmisc)
```
**Dataset and Content**

Results from an experiment to compare yields (as measured by dried
weight of plants) obtained under a control and two different treatment
conditions.Wherein we need to know if there is significant differences
between the means of our different groups.

**Assuption Checks**
Assumption 1:
You have one dependent variable that is measured at the continuous level.
```{r}
head(PlantGrowth)
```
**Remark:**
Given the data shown above our independent variable are the different groups that the experiment has and the weight is the dependent variable where weight is measured at a continous level.

Assumption 2:
You have one independent variable that consists of two categorical, independent groups.
```{r}
unique(PlantGrowth$group)
```
**Remark:**
Given by the values above we can observe that there are 3 distinct values for the group feature where we have "ctrl" the control group, "trt1" the treatment group 1, "trt2" the treatment group 2. Therefore, assumption 2 is satisfied by the data provided.

Assumption 3:
You should have independence of observations.

**Remark:**
Based on the given data there are no direct relationships between the levels within the groups therefore assumption 3 is satisfied.

Assumption 4:
There should be no significant outliers in the three or more groups of your independent variable in terms of the dependent variable.

```{r}
boxplot(weight ~ group, data = PlantGrowth,
        main = "Box Plot of the plant weight for each group",
        xlab = "Group",
        ylab = "Weight",
        col = "lightblue")
```
**Remark:**
We can observe that the group for "trt1" contains a significant outlier, however we have a dilemma wherein we cannot necessaryly remove them since each group has the same number of plants wieghted. To avoid Type I and Type 2 errors we can Winsorize the outlier such that it will have the value equated to the nearest non-outlier value.

```{r}
PlantGrowth[17,"weight"] <- 5.87
boxplot(weight ~ group, data = PlantGrowth,
        main = "Box Plot of the plant weight for each group",
        xlab = "Group",
        ylab = "Weight",
        col = "lightblue")
```
With Winsorize Assumption 4 is satisfied.

Assumption 5:
Your dependent variable should be approximately normally distributed for each group of the independent variable.

```{r}
ShapiroWilkTest <- PlantGrowth %>%
  group_by(group) %>%
  summarise(
    shapiro_statistic = shapiro.test(weight)$statistic,
    shapiro_pValue = shapiro.test(weight)$p.value
  )
ShapiroWilkTest
```
**Remark:**
Since the p-value for each group is normal by our large p-values, therefore the the dependent variable for each group is normally distributed, hence assumption 5 is satisfied.

Assumption 6:
You have homogeneity of variances (i.e., the variance of the dependent variable is equal in each group of your independent variable).
```{r}
LeveneTest <- PlantGrowth %>%
  group_by(group) %>%
  reframe(
    leveneStatistic = leveneTest(weight ~ group, data = .)$`F value`,
    levene_pValue = leveneTest(weight ~ group, data = .)$`Pr(>F)`
  )
LeveneTest
```

**Remark:**
Given by the p-value of the levene's test for each groups we can determine that the variances of the dependent variable are equal across all groups.Therefore homogeneity of variance is true.

**Descriptive Statistics**
```{r}
describe(PlantGrowth$weight)
```

Since our dataset passed all assumptions of the One-way ANOVA we can perform the test.

H_0: There is no significant difference on weight between treatment groups
H_1: There is a significant difference of weight on at least one of the treatment groups
```{r}
anova_result <- aov(weight ~ group, data = PlantGrowth)
summary(anova_result)
```
Given by the p-value we can reject our null hypothesis, therefore we can say that there exist significant difference in means between the three groups.

**Post-Hoc Test**
Perform using tukeys
```{r}
anova_result <- aov(weight ~ group, data = PlantGrowth)
tukey <- TukeyHSD(anova_result)
tukey
```
By the post-hoc test shows that only the comparison between trt2 and trt1 are significant differences whilst the other comparisons are non- significant.
